Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
#Don't need to specify x and y in qplot
qplot(age, friend_count, data = pf)
#Equivalent with ggplot syntax
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point()
```

***

#### What are some things that you notice right away?
Response:
The people with the higher friend counts tends to be younger than 28 years old. There's a strange spike around 110 years old.
***

### ggplot Syntax
Notes:
Instructor recommends adding one layer at a time when building plots, for easier debugging

```{r ggplot Syntax}
summary(pf$age)

#Equivalent with ggplot syntax
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13,90)
```

***

### Overplotting
Notes:
When too many points are overlapping.
```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
#Making each point on the chart equal to 20 data points
      geom_point(alpha = 1/20) +
  xlim(13,90)

#jitter geom adds a small amount of random variation to the location of each point.
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13,90)

```

#### What do you notice in the plot?
Response:
There are more users in general under 25. The data is left skewed.
It is now easier to see that younger users does not have that many more friends than older users.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13,90)
  
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13,90) +
  coord_trans(x = "log10") +
  geom_smooth()
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
#See above
```

#### What do you notice?
40something percent seems to be under 30 years old. 13 year olds have relatively many friends, and then there's a slow decline in friend count to around 20, when the decline becomes more rapid until 30. From 30 to 50 the friend count is pretty stable, and rises slowly from 50 to 80, and at 80 the count is about the same as at 28. From 80 to 90 there's a rapid increase, ending up at the same level as 13 year olds. We can speculate that this is due to some under 20 year olds using a false age of 80+ on their profile.

Answer from video:
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
```
It is now easier to see the distribution of friend count conditional on age. For example, we can see thresholds of friend count above where there are very few users.
***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
```
**QUIZ**
# Examine the relationship between
# friendships_initiated (y) and age (x)
# using the ggplot syntax.

# We recommend creating a basic scatter
# plot first to see what the distribution looks like.
# and then adjusting it by adding one layer at a time.

# What are your observations about your final plot?

# Remember to make adjustments to the breaks
# of the x-axis and to use apply alpha and jitter.

***
```{r}
summary(pf[,c("age","friendships_initiated","friend_count")])

ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0) ) +
  scale_x_continuous(limits = c(13,90), breaks = seq(10,100,10)) +
  scale_y_continuous(breaks = seq(0,4000,100)) +
  coord_trans(y = "sqrt") +
  geom_smooth() +
  ggsave("L5Q7.png",limitsize = TRUE)

#Adding my own, using more interesting variables for comparison
ggplot(aes(x = friendships_initiated, y = friend_count), data = pf) +
  geom_point() +
  scale_x_continuous(limits = c(0,1000),breaks = seq(0,1000,100)) +
  scale_y_continuous(breaks = seq(0,4000,100)) +
  coord_trans(y="sqrt") +
  geom_smooth() 

library(dplyr)
age_13_to_25 <- filter(pf,age<26,age>13)
age_41_to_50 <- filter(pf,age<50,age>40)
age_81_to_90 <- filter(pf,age<90,age>80)

summary(age_13_to_25$friendships_initiated)
summary(age_41_to_50$friendships_initiated)
summary(age_81_to_90$friendships_initiated)
```
The observations here are similar to the scatterplot with age and friend_count: The younger users (about 25 and younger) have both a larger range and a higher average of friendships_initiated than older users. The user group between 80 and 90 years old shows a similar trend. There are less users in this age range, 817, but still far many enough to be statistically significant.
```{r}
#Answer from video
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/10, position = position_jitter(h = 0) ) +
  coord_trans(y = "sqrt")
```

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

filter()

#Grouping by age, adding to new variable
age_groups <- group_by(pf, age)

#summarising friend count for the age groups
pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
#Sorting by age
pf.fc_by_age <- arrange(pf.fc_by_age, age)

#Same as above, using pipe operator
pf.fc_by_age <- group_by(pf, age) %>% 
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>% 
  arrange(age)
  
pf.fc_by_age

```

Create your plot! Plotting mean friend count for the age groups.

```{r Conditional Means Plot}

ggplot() + 
  geom_line(aes(x = age, y = friend_count_mean), data = pf.fc_by_age,color = 'blue') +
  scale_x_continuous(breaks = seq(10,110,5)) +
  scale_y_continuous(breaks = seq(0,500,50))

ggplot(economics_long, aes(date, value01, colour = variable)) +
  geom_line()

ggplot(tmp_bah, aes(age, count, colour = desc_type)) +
  geom_line()

economics_long
#BONUS, for myself:
library(tidyr)
#combining mean and median, sort of. For plotting purposes.
tmp bah <- gather(pf.fc_by_age,desc_type,count,-age,-n)

ggplot(tmp_bah, aes(age, count, colour = desc_type)) +
   geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 0.05, position = position_jitter(h = 0), color = 'orange' ) +
  scale_x_continuous(limits = c(13,90), breaks = seq(10,100,10)) +
  scale_y_continuous(breaks = seq(0,4000,100)) +
  coord_trans(y = "sqrt") + 
  #adding line.
  geom_line(stat = "summary", fun.y = mean) + 
  #adding quantile for bottom 10%
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .1), 
            linetype = 2, color = 'blue') +
  #adding quantile for bottom 90%
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .9), 
            linetype = 2, color = 'black') +
  #adding quantile for bottom 50% (AKA median)
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .5), 
            linetype = 1, color = 'blue')
  
#Gives us some more knowledge, but the spread is still unclear

```

#### What are some of your observations of the plot?
Response:
The mean is consistently quite higher than the median for all age groups, indicating the presence of some large outliers. 

Observations from answer:
Having more than 1000 friends is quite rare, including for younger users. the 9- percentile is never above 500.
```{r}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 0.05, position = position_jitter(h = 0), color = 'orange' ) +
  #adding coord cartesian
  coord_cartesian(xlim = c(13,70),ylim = c(0,1000)) +
  geom_line(stat = "summary", fun.y = mean) + 
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .1), 
            linetype = 2, color = 'blue') +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .9), 
            linetype = 2, color = 'black') +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .5), 
            linetype = 1, color = 'blue')
```

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age,pf$friend_count)

#alternative way
with(pf, cor.test(age, friend_count))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.0274
***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age >= 13 & age <=70 & 
              friend_count <= 1000 & friend_count >= 0), cor.test(age, friend_count))

with(subset(pf, age <=70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:
Plots give a richer understanding of correlation than what statistical tests (r, rho, tau) do.
***

## Create Scatterplots
Notes:
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.

```{r}
quantile(pf$likes_received, probs = c(0.5,0.9,0.95,0.99))

#Looking at 90th percentile
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(color = 'orange') +
  scale_y_continuous(limits = c(0,600)) +
  scale_x_continuous(limits = c(0,600)) +
  geom_line(stat = 'summary', fun.y = mean)

#From solution video
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(color = 'orange')
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) + 
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:
Pearson's r is 0.948.

"The correlation coefficient is invariant under a linear transformation of either X or Y, and the slope of the regression line when both X and Y have been transformed to z-scores is the correlation coefficient.

It's important to note that we may not always be interested in the bulk of the data. Sometimes, the outliers ARE of interest, and it's important that we understand their values and why they appear in the data set."
***

### Moira on Correlation
Notes:
Strong correlations is not always a good thing. Sometime this just indicates they are measuring the same thing (e.g. engagement n to Facebook).
"When working on a problem, and I am doing some kind of regression or I am modelling something, I will be throwing some of these variable into the regression. One of the assumptions of the regression is that the variables are independent of eachother. If any two are highly correlated with eachother, it will be very difficult to tell which one is actually driving the phenomenon. So it is important to measure the correlation between your variables first, often because it will help you determine which ones you don't want to throw in together, and it will help you figure out which ones you actually want to keep."

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data("Mitchell")

?Mitchell

Mitchell
```

Create your plot!
# Create a scatterplot of temperature (Temp)
# vs. months (Month).
```{r Temp vs Month}
head(Mitchell)

ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0.1

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
#Getting info about the dataframe
?Mitchell

ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  geom_smooth(method = 'lm')

cor.test(Mitchell$Month,Mitchell$Temp)


```
0.057
***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
#Getting info about dataset
?Mitchell
#Adding month name to Mitchell data
Mitchell$Month_name <- month.name


ggplot(aes(x = match(Mitchell$Month_name,month.name), y = Temp), data = Mitchell) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  scale_x_continuous(breaks = seq(1,12,1))

cor.test(match(Mitchell$Month_name,month.name),Mitchell$Temp)
```

```{r}
years_list = c()
#Creating list of years
for (i in seq(1976,1975+(nrow(Mitchell)/12),1)) {
  for (mo in month.name) {
    #print(i)
    years_list <- c(years_list,i)
  }
}
#Adding years to Mitchell data
Mitchell$year <- years_list
#Making year column type categorical, for plotting purposes
Mitchell$year <- factor(Mitchell$year)

ggplot(aes(x = match(Mitchell$Month_name,month.name), y = Temp, color = fyear), data = Mitchell) +
  geom_point() +
#  geom_line() +
  scale_x_continuous(breaks = seq(1,12,1))

cor.test(Mitchell$Temp,as.numeric(Mitchell$year))

?scale_x_continuous()
```
I misunderstood the question a little, I was supposed to solve this within the scale_x_continuos() scale. My work is much more thorough and insightful.

```{r}
#Answer from video
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  geom_smooth(method = 'lm') +
  scale_x_continuous(breaks = seq(0,203,12))

```


***

### A New Perspective
```{r}
#Alternative method
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()

```

What do you notice?
Response:
The distribution is now normal. Temperatures tend to be similar in the same month from year to year. Temperatures follow the seasons.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:
# Create a new variable, 'age_with_months', in the 'pf' data frame.
# Be sure to save the variable in the data frame rather than creating
# a separate, stand-alone variable. You will need to use the variables
# 'age' and 'dob_month' to create the variable 'age_with_months'.

# Assume the reference date for calculating age is December 31, 2013.

```{r Understanding Noise: Age to Age Months}
View(pf)
#Date december 31
#month_decimals <- ifelse(pf$dob_month == 12 & pf$dob_day == 31,0,round( (12 - pf$dob_month) /12,2) )

pf$age_with_months <- pf$age + round( (12 - pf$dob_month) /12,2)

#The grader does not accept including round function. Removed here:
pf$age_with_months <- pf$age + (12 - pf$dob_month) /12


ggplot(aes(x = age_with_months, y = friend_count), data = pf) + 
  geom_line(stat = 'summary', fun.y = mean)

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_line(stat = 'summary', fun.y = mean)
```

***

### Age with Months Means

# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data framed should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11
```{r Age with Months Means}
#library(dplyr)
pf.fc_by_age_months <- pf %>% group_by(age_with_months) %>% 
  summarise(friend_count_mean = mean(friend_count), friend_count_median = 
              median(friend_count), n = n()) %>% 
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means
# Create a new line plot showing friend_count_mean versus the new variable,
# age_with_months. Be sure to use the correct data frame (the one you created
# in the last exercise) AND subset the data to investigate users with ages less
# than 71.
```{r Noise in Conditional Means}

ggplot(data = pf.fc_by_age_months, aes(x = age_with_months, y = friend_count_mean) ) +
  geom_line() + 
  scale_x_continuous(limits = c(13,70), breaks = seq(10,70,5) ) + 
  scale_y_continuous(limits = c(0,450) )

#Solution from video (with minor adjustments), both work
ggplot(data = subset(pf.fc_by_age_months, age_with_months < 71),
       aes(x = age_with_months, y = friend_count_mean) ) +
  geom_line() + 
  scale_x_continuous(limits = c(13,70), breaks = seq(10,70,5) ) + 
  scale_y_continuous(limits = c(0,450) )

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(data = subset(pf.fc_by_age, age < 71),
       aes(x = age, y = friend_count_mean) ) +
  geom_line()

p2 <- ggplot(data = subset(pf.fc_by_age_months, age_with_months < 71),
       aes(x = age_with_months, y = friend_count_mean) ) +
  geom_line()

library(gridExtra)

p3 <- ggplot(data = subset(pf, age < 71),
       aes(x = round(age / 5) * 5, y = friend_count) ) +
  geom_line(stat = 'summary', fun.y = mean)


grid.arrange(p2, p1, p3, ncol = 1)
```
Running this, we see an example of bias variance tradeoff
- Less data points and wider bin widths 
- We're calculating mean more precicely, but potentially miss out on important features of the age and friend count relationship
- Similar to what we see when we choose bin widths in histograms
- To improve this tradeoff, analysts can use a flexible statistical model, to smooth our estimates of conditional means. In ggplot, we use geom_smooth() for this
```{r}
#Same as above, using smooth()
p1 <- ggplot(data = subset(pf.fc_by_age, age < 71),
       aes(x = age, y = friend_count_mean) ) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(data = subset(pf.fc_by_age_months, age_with_months < 71),
       aes(x = age_with_months, y = friend_count_mean) ) +
  geom_line() +
  geom_smooth()

grid.arrange(p2, p1, p3, ncol = 1)
```
For p1, the smoother does not draw attention to the "non-matonic relationship" in the low ages, and it really misses the discontinuity at age 69. This highlights that using models like Loess or smoothing lines can be useful, but like almost any model can be subject to systematic errors when the true process generating our data is not so consistent with the model itself.
***

### Which Plot to Choose?
Notes:
For exploratory analysis, you don't need to choose one plot type. Part of the process is to visualize the data in different ways. But when it comes to presenting your findings and/or the data, one or two visualizations should be chosen.
***

### Analyzing Two Variables
Reflection:
I have learned more about scatter plots, and how to create them with various features using ggplot(). I have gotten to use dplyr more for data wrangling, and have touched on using colors as a way to visualize 3-dimensions on a plot. Sometimes to make sense of data, the data first have to be transformed into different units (e.g. month number in experiment converted to month of the year).

From instructor:
We learned how to explore the relationship between two variables
Learned about benefits and limitations of using correlation to understand the relationship between two variables. 
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

