Lesson 5
========================================================

### Multivariate Data
Notes:
```{r}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

***

### Moira Perceived Audience Size Colored by Age
Notes:
Using color to represent age of survey respondents (third variable)
***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() + 
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

#Comparing median friend count for female and males of different ages
ggplot(data = subset(pf, !is.na(gender)), aes(x = age, y = friend_count)) +
  geom_line(stat = 'summary', fun.y = median, aes(color = gender))
```
QUIZ:
```{r}
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

library(dplyr)

pf.fc_by_age_gender <- group_by(pf, age, gender) %>% 
  summarise(mean_friend_count = mean(friend_count), 
            median_friend_count = median(friend_count),
            n = n())

#This is mostly correct, but I forgot to remove null values.
#Rhe solution from the video is like this:
pf.fc_by_age_gender <- pf %>% 
  filter(!is.na(gender)) %>% #to remove null values
  group_by(age, gender) %>% 
  summarise(mean_friend_count = mean(friend_count), 
            median_friend_count = median(friend_count),
            n = n()) %>% 
  ungroup() %>% #ungrouping to use arrange
  arrange(age)

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(data = pf.fc_by_age_gender, 
       aes(x = age, y = median_friend_count, color = gender)) +
  geom_line()

#From solution video, gives same result
ggplot(data = pf.fc_by_age_gender, 
       aes(x = age, y = median_friend_count)) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:
How many times more friends do females have?
***

### Wide and Long Format
Notes:
"Many times when computing with and exploring data it is helpful to move back and forth between 
these different arrangements". To carry this out in r, we can use the reshape2 package.
***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

#dcast() returns dataframe
#To return array or matrix, use acast()

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender, #left of ~ we keep, ~right what we change
                                  value.var = 'median_friend_count') #column to take values from

pf.fc_by_age_gender.wide
pf.fc_by_age_gender
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

head(pf.fc_by_age_gender.wide)

ggplot(data = pf.fc_by_age_gender.wide, aes(x = age, y = female/male)) + 
  geom_point() +
  geom_smooth() +
  geom_hline(yintercept = 1, linetype = 2)

#Solution from video
ggplot(data = pf.fc_by_age_gender.wide, aes(x = age, y = female / male)) + 
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2) #alpha to make the line more transparent
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

pf$year_joined <-  as.numeric(substring(2014 - (pf$tenure / 365),1,4))

#Solution from video (better)
pf$year_joined <-  floor(2014 - (pf$tenure / 365))
#Opposite of floor(): ceiling()
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)

# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

?cut

pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks = c(2004,2009,2011,2012,2014))
#See https://chrisalbon.com/r-stats/cutting-up-data.html for simple tutorial on cut()
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
#table() gives counts at each combination of factor levels
table(pf$year_joined.bucket, useNA = 'ifany')

# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

ggplot(data = subset(pf, !is.na(year_joined.bucket)), 
       aes(x = age, y = friend_count, color = year_joined.bucket)) +
  geom_line(stat = 'summary', fun.y = median)

#Solution from video
ggplot(data = subset(pf, !is.na(year_joined.bucket)), 
       aes(x = age, y = friend_count, color = year_joined.bucket)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

ggplot(data = subset(pf, !is.na(year_joined.bucket)), 
       aes(x = age, y = friend_count, color = year_joined.bucket)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(aes(x = age, y = friend_count), stat = 'summary', fun.y = mean,
            linetype = 2, color = 'black')

#Solution from video. I overcomplicated it above (same result though)
ggplot(data = subset(pf, !is.na(year_joined.bucket)), 
       aes(x = age, y = friend_count, color = year_joined.bucket)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean,
            linetype = 2, color = 'black')

```
"Plotting the grand mean is a good reminder that much of the data in the sample is about members of the recent cohort. This is a high-level observation you want to make as you explore the data."
***

### Friending Rate
Notes:

```{r Friending Rate}
pf.tenure_above_zero = subset(pf, tenure > 0)

pf.tenure_above_zero$tenure

pf.tenure_above_zero$friends_per_day <- pf.tenure_above_zero$friend_count / pf.tenure_above_zero$tenure

summary(pf.tenure_above_zero$friends_per_day)

#Solution from video. More elegant, same answer.
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```
Median friend rate: 0.22
Max friend rate: 417


***

### Friendships Initiated
Notes:

What is the median friend rate?
0.22
What is the maximum friend rate?
417
```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

ggplot(data = pf.tenure_above_zero, aes(x = tenure, y = friends_per_day,
                                        color = year_joined.bucket)) +
  geom_line(stat = 'summary', fun.y = mean)
  
ggplot(data = subset(pf.tenure_above_zero, tenure > 100),
                     aes(x = tenure, y = friends_per_day,
                                        color = year_joined.bucket)) +
  geom_line(stat = 'summary', fun.y = mean)

#Solution from video
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

library(gridExtra)

p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

grid.arrange(p1,p2,p3,p4, ncol = 1)
# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

```
NOTE: The code changing the binning is substituting x = tenure in the plotting expressions with x = 7 * round(tenure / 7), etc., binning values by the denominator in the round function and then transforming back to the natural scale with the constant in front.
For more, see http://scott.fortmann-roe.com/docs/BiasVariance.html
***

### Sean's NFL Fan Sentiment Study
Notes:
Expanding the time periods and adding spine(?) makes the data smoother/ trends clearer.
***

### Introducing the Yogurt Data Set
Notes:
Household purchases of 5 flavors of yougurt 8 oz size with price.

Many rows per househould, one line per purchase.

```{r}
yo = read.csv('yogurt.csv')

head(yo)
str(yo)

#Making id into categorical variable
yo$id <- factor(yo$id)
str(yo)

```

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
ggplot(data = yo, aes(x = price)) +
  geom_histogram(binwidth = 5) + 
  scale_x_continuous(breaks = seq(0,80,5))

ggplot(data = yo, aes(x = price)) +
  geom_histogram()
```
Some price points are much more common than others.


***

### Number of Purchases
Notes:

```{r Number of Purchases}


summary(yo)

#We see that there's only 20 unique price points
length(unique(yo$price))

# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

yo$all.purchases

?transform

yo <- transform(yo, 
          all.purchases = strawberry + blueberry + pina.colada +
            plain + mixed.berry)

head(yo)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(data = yo, aes(x = all.purchases)) +
  geom_histogram(fill = 'lightblue')
```

```{r Prices over Time}
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?

ggplot(data = yo, aes(y = price, x = time)) +
  geom_point(alpha = 0.1) + 
  geom_smooth()

#Solution from video
ggplot(aes(x  = time, y = price), data = yo) + 
  geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))

```
I notice that in general the prices tend to be fairly high with little variation from day to day. The modal price tend to go up over time.

***

### Sampling Observations
Notes:
When familiarizing yourself with a new dataset containing multiple observations of the same unit, it is often useful to work with a sample of those units, so that it is easy to display the raw data for that sample.
***

### Looking at Samples of Households

```{r Looking at Sample of Households}
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?

ggplot(data = yo, aes(y = price, x = time)) +
  geom_point(alpha = 0.1) + 
  geom_smooth()

#Solution from video
ggplot(aes(x  = time, y = price), data = yo) + 
  geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))

```
I notice that in general the prices tend to be fairly high with little variation from day to day. The modal price tend to go up over time.


***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

