Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
#Don't need to specify x and y in qplot
qplot(age, friend_count, data = pf)
#Equivalent with ggplot syntax
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point()
```

***

#### What are some things that you notice right away?
Response:
The people with the higher friend counts tends to be younger than 28 years old. There's a strange spike around 110 years old.
***

### ggplot Syntax
Notes:
Instructor recommends adding one layer at a time when building plots, for easier debugging

```{r ggplot Syntax}
summary(pf$age)

#Equivalent with ggplot syntax
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13,90)
```

***

### Overplotting
Notes:
When too many points are overlapping.
```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
#Making each point on the chart equal to 20 data points
      geom_point(alpha = 1/20) +
  xlim(13,90)

#jitter geom adds a small amount of random variation to the location of each point.
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13,90)

```

#### What do you notice in the plot?
Response:
There are more users in general under 25. The data is left skewed.
It is now easier to see that younger users does not have that many more friends than older users.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13,90)
  
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13,90) +
  coord_trans(x = "log10") +
  geom_smooth()
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
#See above
```

#### What do you notice?
40something percent seems to be under 30 years old. 13 year olds have relatively many friends, and then there's a slow decline in friend count to around 20, when the decline becomes more rapid until 30. From 30 to 50 the friend count is pretty stable, and rises slowly from 50 to 80, and at 80 the count is about the same as at 28. From 80 to 90 there's a rapid increase, ending up at the same level as 13 year olds. We can speculate that this is due to some under 20 year olds using a false age of 80+ on their profile.

Answer from video:
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
```
It is now easier to see the distribution of friend count conditional on age. For example, we can see thresholds of friend count above where there are very few users.
***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
```
**QUIZ**
# Examine the relationship between
# friendships_initiated (y) and age (x)
# using the ggplot syntax.

# We recommend creating a basic scatter
# plot first to see what the distribution looks like.
# and then adjusting it by adding one layer at a time.

# What are your observations about your final plot?

# Remember to make adjustments to the breaks
# of the x-axis and to use apply alpha and jitter.

***
```{r}
summary(pf[,c("age","friendships_initiated","friend_count")])

ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0) ) +
  scale_x_continuous(limits = c(13,90), breaks = seq(10,100,10)) +
  scale_y_continuous(breaks = seq(0,4000,100)) +
  coord_trans(y = "sqrt") +
  geom_smooth() +
  ggsave("L5Q7.png",limitsize = TRUE)

#Adding my own, using more interesting variables for comparison
ggplot(aes(x = friendships_initiated, y = friend_count), data = pf) +
  geom_point() +
  scale_x_continuous(limits = c(0,1000),breaks = seq(0,1000,100)) +
  scale_y_continuous(breaks = seq(0,4000,100)) +
  coord_trans(y="sqrt") +
  geom_smooth() 

library(dplyr)
age_13_to_25 <- filter(pf,age<26,age>13)
age_41_to_50 <- filter(pf,age<50,age>40)
age_81_to_90 <- filter(pf,age<90,age>80)

summary(age_13_to_25$friendships_initiated)
summary(age_41_to_50$friendships_initiated)
summary(age_81_to_90$friendships_initiated)
```
The observations here are similar to the scatterplot with age and friend_count: The younger users (about 25 and younger) have both a larger range and a higher average of friendships_initiated than older users. The user group between 80 and 90 years old shows a similar trend. There are less users in this age range, 817, but still far many enough to be statistically significant.
```{r}
#Answer from video
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/10, position = position_jitter(h = 0) ) +
  coord_trans(y = "sqrt")
```

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

filter()

#Grouping by age, adding to new variable
age_groups <- group_by(pf, age)

#summarising friend count for the age groups
pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
#Sorting by age
pf.fc_by_age <- arrange(pf.fc_by_age, age)

#Same as above, using pipe operator
pf.fc_by_age <- group_by(pf, age) %>% 
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>% 
  arrange(age)
  
pf.fc_by_age

```

Create your plot! Plotting mean friend count for the age groups.

```{r Conditional Means Plot}

ggplot() + 
  geom_line(aes(x = age, y = friend_count_mean), data = pf.fc_by_age,color = 'blue') +
  scale_x_continuous(breaks = seq(10,110,5)) +
  scale_y_continuous(breaks = seq(0,500,50))

ggplot(economics_long, aes(date, value01, colour = variable)) +
  geom_line()

ggplot(tmp_bah, aes(age, count, colour = desc_type)) +
  geom_line()

economics_long
#BONUS, for myself:
library(tidyr)
#combining mean and median, sort of. For plotting purposes.
tmp bah <- gather(pf.fc_by_age,desc_type,count,-age,-n)

ggplot(tmp_bah, aes(age, count, colour = desc_type)) +
+   geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 0.05, position = position_jitter(h = 0), color = 'orange' ) +
  scale_x_continuous(limits = c(13,90), breaks = seq(10,100,10)) +
  scale_y_continuous(breaks = seq(0,4000,100)) +
  coord_trans(y = "sqrt") + 
  #adding line.
  geom_line(stat = "summary", fun.y = mean) + 
  #adding quantile for bottom 10%
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .1), 
            linetype = 2, color = 'blue') +
  #adding quantile for bottom 90%
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .9), 
            linetype = 2, color = 'black') +
  #adding quantile for bottom 50% (AKA median)
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .5), 
            linetype = 1, color = 'blue')
  
#Gives us some more knowledge, but the spread is still unclear

```

#### What are some of your observations of the plot?
Response:
The mean is consistently quite higher than the median for all age groups, indicating the presence of some large outliers. 

Observations from answer:
Having more than 1000 friends is quite rare, including for younger users. the 9- percentile is never above 500.
```{r}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 0.05, position = position_jitter(h = 0), color = 'orange' ) +
  #adding coord cartesian
  coord_cartesian(xlim = c(13,70),ylim = c(0,1000)) +
  geom_line(stat = "summary", fun.y = mean) + 
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .1), 
            linetype = 2, color = 'blue') +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .9), 
            linetype = 2, color = 'black') +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .5), 
            linetype = 1, color = 'blue')
```

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age,pf$friend_count)

#alternative way
with(pf, cor.test(age, friend_count))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.0274
***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age >= 13 & age <=70 & 
              friend_count <= 1000 & friend_count >= 0), cor.test(age, friend_count))

with(subset(pf, age <=70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:
Plots give a richer understanding of correlation than what statistical tests (r, rho, tau) do.
***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(color = 'orange') +
  geom_line(stat = 'summary', fun.y = mean)

ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 0.05, position = position_jitter(h = 0), color = 'orange' ) +
  #adding coord cartesian
  #coord_cartesian(xlim = c(13,70),ylim = c(0,1000)) +
  geom_line(stat = "summary", fun.y = mean)

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

