---
title: "R Notebook"
output: html_notebook
---

*1. Quiz: price vs. x*
# In this problem set, you'll continue
# to explore the diamonds data set.

# Your first task is to create a
# scatterplot of price vs x.
# using the ggplot syntax.
```{r}
head(diamonds)
ggplot(data = diamonds, aes(x = x, y = price)) + 
  geom_point(color = 'orange') + 
  geom_line(stat = 'summary', fun.y = mean)
```

*2. Quiz: Findings - price vs. x*
Observations:
The releationship seems quite linear. Although the price range for x is quite wide, price does steadily go up as x rises.
From answer:
"Did you notice some outliers and an exponential relationship between price and x?
"

*3. Quiz: Correlations*
What is the correlation between price and x?
```{r}
cor.test(diamonds$price, diamonds$x)
```
0.88
What is the correlation between price and y?
```{r}
cor.test(diamonds$price, diamonds$y)
```
0.87
What is the correlation between price and z?
```{r}
cor.test(diamonds$price, diamonds$z)
```
0.86

*4. Quiz: price vs. depth*
# Create a simple scatter plot of price vs depth.
```{r}
ggplot() + 
  geom_point(data = diamonds, aes(x = depth, y = price))
```

*5. Quiz: Adjustments - price vs. depth*
# Change the code to make the transparency of the
# points to be 1/100 of what they are now and mark
# the x-axis every 2 units. See the instructor notes
# for two hints.
```{r}
summary(diamonds$depth)

ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point(alpha = 1/100) + 
  scale_x_continuous(breaks = seq(40,80,2))
```

*6. Quiz: Typical depth range*
Based on the scatterplot of depth vs. price, most diamonds are between what values of depth?
59 to 64

Control (own initiative):

```{r}
library(dplyr)

View(group_by(diamonds,round(depth,0)) %>% 
  summarise(count = n()))

```

*7. Quiz: Correlation - price and depth*
What's the correlation of depth vs. price?
```{r}
cor.test(diamonds$depth, diamonds$price)
```
-0.01
Based on the correlation coefficient would you use depth to predict the price of a diamond? Why?
No. Because the according to the correlation coefficient there is no statistically significant relationship between the two variables.

*8. Quiz: price vs. carat*
# Create a scatterplot of price vs carat
# and omit the top 1% of price and carat
# values.
```{r}
#notes
quantile(diamonds$price,0.99)
nrow(diamonds)
nrow(subset(diamonds, price < quantile(diamonds$price,0.99) &
                     carat < quantile(diamonds$carat,0.99)))

#the plot
ggplot(data = subset(diamonds, price < quantile(diamonds$price,0.99) &
                     carat < quantile(diamonds$carat,0.99)),
       aes(x = carat, y = price) ) + 
  geom_point()

```

*9. Quiz: price vs. volume*
# Create a scatterplot of price vs. volume (x * y * z).
# This is a very rough approximation for a diamond's volume.
# Create a new variable for volume in the diamonds data frame.
# This will be useful in a later exercise.
# Don't make any adjustments to the plot just yet.
```{r}
ggplot(data = diamonds, aes(x = (x*y*z), y = price)) + 
  geom_point()

diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(data = diamonds, aes(x = volume, y = price)) + 
  geom_point()

subset(diamonds, volume == 0)
```

*10. Quiz: Findings - price vs volume*
Observations:
The general trend is that price goes sharply upwards as volume increases.
There's a few noticable outliers, where the volume is much higher per price point than the general trend. There are also quite a few diamonds with volume 0, which must be due to lack of data for at least one of the x, y and z variables.
```{r}
#To get count
library(plyr)
count(diamonds$volume == 0)
#Detaching plyr to avoid conflicts with dplyr
detach("package:plyr", unload = TRUE)
```

*11. Quiz: Correlations on subset*
What's the correlation on volume? Exclude diamonds that have a volume of 0 or that are greater than or equal to 800.
```{r}
vol_diamonds <- subset(diamonds, volume > 0 & volume <= 800)

cor.test(vol_diamonds$price, vol_diamonds$volume)
```
0.92

*12. Quiz: Adjustments - price vs. volume*
# Subset the data to exclude diamonds with a volume
# greater than or equal to 800. Also, exclude diamonds
# with a volume of 0. Adjust the transparency of the
# points and add a linear model to the plot. (See the
# Instructor Notes or look up the documentation of
# geom_smooth() for more details about smoothers.)

```{r}
ggplot(data = vol_diamonds, aes(y = price, x = volume)) + 
  geom_point(alpha = 1/20) + 
  geom_smooth(method = 'lm') +
  ggsave('pricevsvolume.png')
```
# Do you think this would be a useful model to estimate
# the price of diamonds? Why or why not?
Partially useful. There is definitely a strong correlation between volume and price of the diamonds, but other factors also have a large impact. For example, when we use color to visualize diamond cut, we see that a Fair cut diamond has to be about 25% larger than an Ideal cut diamond before it reaches the 10 000 price point.

```{r}
ggplot(data = vol_diamonds, aes(y = price, x = volume, color = cut)) + 
  geom_point(alpha = 1/20) + 
  geom_smooth(method = 'lm') +
  ggsave('pricevsvolume2.png')
```
The relationship between price and volume does not appear to be strictly linear, and an exponential model might be a better fit.  
```{r}

ggplot(data = vol_diamonds, aes(y = price, x = volume, color = color)) + 
  geom_point(alpha = 1/20) + 
  geom_smooth(method = 'lm')
```

*13. Quiz: Mean price by clarity*
# Use the function dplyr package
# to create a new data frame containing
# info on diamonds by clarity.

# Name the data frame diamondsByClarity

# The data frame should contain the following
# variables in this order.

#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

# where n is the number of diamonds in each
# level of clarity.
```{r}
library(dplyr)

diamondsByClarity <- group_by(diamonds,clarity) %>% 
  summarise(mean_price = mean(price), median_price = median(price), 
            min_price = min(price), max_price = max(price), n = n())

diamondsByClarity  
```

*14. Quiz: Bar Charts of Mean Price*
```{r}
#Provided by quiz description
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```
# Your task is to write additional code to create two bar plots
# on one output image using the grid.arrange() function from the package
# gridExtra.
```{r}
library(gridExtra)
diamonds_mp_by_clarity
diamonds_mp_by_color

p1 <- ggplot(data = diamonds_mp_by_clarity,aes(x = clarity, y = mean_price)) +
  geom_col()
p2 <- ggplot(data = diamonds_mp_by_color,aes(x = color, y = mean_price)) +
  geom_col()

grid.arrange(p1, p2, ncol = 2)
```

*15. Quiz: Trends in Mean Price*
What do you notice in each of the bar charts for mean price by clarity and mean price by color?
For color, mean price goes up the poorer the color quality gets; the graph is right-skewed. For clarity, the mean price is fairly stable until clarity VVS2, except for SI2, which is more than 20% higher than I1, SI1, VS2 and VS1. 

From solution:
"We think something odd is going here. These trends seem to go against our intuition.
Mean price tends to decrease as clarity improves. The same can be said for color.
We encourage you to look into the mean price across cut."
```{r}
diamonds_mp_by_cut <- group_by(diamonds, cut) %>% 
  summarise(mean_price = mean(price))

p3 <- ggplot(data = diamonds_mp_by_cut,aes(x = cut, y = mean_price)) +
  geom_col()

grid.arrange(p1, p2, p3, nrow = 1)
```

*16. Quiz: Gapminder revisited*
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to continue the investigation you did at the
# end of Problem Set 3 or you can start fresh and choose a different
# data set from Gapminder.

# If you’re feeling adventurous or want to try some data munging see if you can
# find a data set or scrape one from the web.

# In your investigation, examine pairs of variable and create 2-5 plots that make
# use of the techniques from Lesson 4.

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. the variable(s) you investigated, your observations, and any summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots
```{r}
#THIS CELL NEEDS CLEANUP (TBD)

library(ggplot2)
#Loading data
fertility <- read.csv('indicator undata total_fertility.csv', header = T, row.names = 1, check.names = F)
fertility2 <- read.csv('indicator undata total_fertility.csv', header = T, row.names = 1)

fertility2
#install.packages('tibble')
library(tibble)

#Adding column for Country, for easier wrangling
fertility <- rownames_to_column(fertility, var = "country")
fertility2 <- rownames_to_column(fertility2, var = "country")

#Getting initial overview of fertility rates per country in 2015
ggplot(aes(x = country, y = fertility$'2015'), data = fertility) + 
  geom_point()

fertility$'2015'
qplot(fertility$'2015', data = fertility)
summary(fertility$'2015')

View(fertility[-3])


#Mean per year for all countries
fertility_mean_by_year = data.frame(mean_birth_rate = colMeans(fertility[,-1], na.rm = T))


#Looking for trends per year, sumarized mean for all countries using stack()
ggplot(aes(x = ind, y = values), data = stack(fertility[,-1])) + 
  geom_boxplot() + 
  scale_y_continuous()

library(tidyr)

fertility_by_year <- gather(fertility, key = 'year', value = births, -country)

fertility_mean_by_year <- group_by(fertility_by_year,year) %>% 
  summarise(mean_birth_rate = max(births, na.rm = TRUE))

summary(fertility$'2015')[5] - summary(fertility$'2015')[2]

summary(fertility$'2015')[5,]
summary(fertility$'2015')[2,]

3.495-1.785

ggplot(stack(fertility[200:215]), aes(x = ind, y = values)) + 
  geom_boxplot()

install.packages("countrycode")
library(countrycode)

countrycode(c('norway','usa'),'country.name','region')

countrycode(row.names(fertility),'country.name','region')

fertility$continent <- countrycode(fertility$country,'country.name','continent')
fertility$region <- countrycode(fertility$country,'country.name','region')

fertility[(length(fertility)-2):length(fertility)]

View(tail(countrycode_data))


distinct(countrycode_data,eurocontrol_pru)
distinct(countrycode_data,region)
distinct(countrycode_data,continent)

ggplot(data = fertility, aes(x = region, y = fertility$'2015')) + 
  geom_boxplot()

ggplot(data = fertility, aes(x = continent, y = fertility$'2015')) + 
  geom_boxplot()

ggplot(data = fertility, aes(x = continent, y = fertility$'2015')) + 
  geom_point()

ggplot(data = fertility, aes(x = fertility$'2015')) + 
  geom_histogram()

subset(fertility, region == 'Central America')$country

subset(fertility, country %in% c('Cuba', 'United States', 'Mexico'))$region

unique(subset(fertility, continent == 'Americas')$region)

#Dividing continent "Americas" into North and South
fertility$continent[fertility$region == 'South America'] <- 'South America'
fertility$continent[fertility$continent == 'Americas'] <- 'North America'

#View(fertility[,c('country','continent','region','2015','2014','2010','2000')])

#checking for same length
nrow(fertility)
nrow(fertility2)

#spot checks
fertility[42,1]
fertility2[42,1]
fertility[1,1]
fertility2[1,1]
tail(fertility[,1])
tail(fertility2[,1])

#Adding info about
fertility2$continent <- fertility$continent
fertility2$region <- fertility$region

fertility_mean_by_continent <- group_by(fertility, continent) %>% 
  summarise(mean_2015 = mean(fertility$'2015', na.rm = TRUE), mean_2010 = mean(fertility$'2010'),
            mean_2005 = mean(fertility$'2005'), mean_2000 = mean(fertility$'2000'), 
            mean_1995 = mean(fertility$'1995'), mean_1990 = mean(fertility$'1990'), 
            mean_1985 = mean(fertility$'1985'), mean_1980 = mean(fertility$'1980'), 
            mean_1975 = mean(fertility$'1975'), mean_1970 = mean(fertility$'1970'), 
            mean_1965 = mean(fertility$'1965'), mean_1960 = mean(fertility$'1960'), 
            mean_1955 = mean(fertility$'1955'), mean_1950 = mean(fertility$'1950'), 
            mean_1945 = mean(fertility$'1945'))

fertility_mean_by_continent <- dplyr::group_by(fertility2, continent) %>% 
  dplyr::summarise(y2015 = mean(X2015, na.rm = TRUE), y2010 = mean(X2010, na.rm = TRUE),
            y2005 = mean(X2005, na.rm = TRUE), y2000 = mean(X2000, na.rm = TRUE), 
            y1995 = mean(X1995, na.rm = TRUE), y1990 = mean(X1990, na.rm = TRUE), 
            y1985 = mean(X1985, na.rm = TRUE), y1980 = mean(X1980, na.rm = TRUE), 
            y1975 = mean(X1975, na.rm = TRUE), y1970 = mean(X1970, na.rm = TRUE), 
            y1965 = mean(X1965, na.rm = TRUE), y1960 = mean(X1960, na.rm = TRUE), 
            y1955 = mean(X1955, na.rm = TRUE), y1950 = mean(X1950, na.rm = TRUE), 
            y1945 = mean(X1945, na.rm = TRUE))

#fertility_mean_by_continent <- fertility_mean_by_continent[,order(colnames(fertility_mean_by_continent))]
library(reshape2)
fertility_mean_by_continent <- melt(fertility_mean_by_continent, id.vars = 'continent',
                                    variable.name = 'mean_year', value.name = 'mean_birth_rate')
fertility_mean_by_continent$mean_year <- as.numeric(substr(fertility_mean_by_continent[,2],2,5))

fertility_mean_by_continent

ggplot(data = fertility_mean_by_continent, 
       aes(x = mean_year, y = mean_birth_rate, color = continent)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks = seq(1945,2015,5))


install.packages('wpp2015')
library(wpp2015)

```

