---
title: "R Notebook"
output: html_notebook
---

*Quiz 1: Price Histograms with Facet and Color*
```{r}
# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

# This assignment is not graded and
# will be marked as correct when you submit.
library(ggplot2)
library(gridExtra)

names(diamonds)

ggplot(data = diamonds, aes(x = price, fill = cut)) + 
  geom_histogram() + 
  facet_wrap(~ color)

#Adjusting for positive skew
ggplot(data = diamonds, aes(x = price, fill = cut)) + 
  geom_histogram() + 
  facet_wrap(~ color) +
  coord_trans(x = "log10") + 
  xlim(100,20000)

```

*Quiz 2: Price vs. Table Colored by Cut*
```{r}
# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

head(diamonds)

ggplot(data = diamonds, aes(x = table, y = price, color = cut)) +
  geom_point() +
  scale_color_brewer(type = 'qual') +
  scale_x_continuous(breaks = seq(50,80,2))

```

*Quiz 3: Typical Table Value*
Use the graph from the last exercise.
What is the typical table range for the majority of diamonds of ideal cut?
53 to 57
What is the typical table range for the majority of diamonds of premium cut?
58 to 62
*Quiz 4: Price vs. Volume and Diamond Clarity*
```{r}
# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')

ggplot(data = subset(diamonds,quantile(x*y*z,0.99) > x*y*z), 
       aes(x = (x*y*z), y = price,  color = clarity)) +
  geom_point() +
  scale_y_log10() +
  scale_color_brewer(type = 'div')

?coord_trans()

```

*Quiz 5: Proportion of Friendships Initiated*
```{r}
# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.
pf <- read.table('pseudo_facebook.tsv', header = 1)
head(pf)

pf <- transform(pf,prop_initiated = friendships_initiated / friend_count)
```

*Quiz 6: prop_initiated vs. tenure*
```{r}
# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg

pf$year_joined <- floor(2014 - (pf$tenure / 365))
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))

ggplot(data = pf, aes(x = tenure, y = prop_initiated, color = year_joined.bucket)) +
  geom_line(stat = 'summary', fun.y = median)
```

*Quiz 7: smoothing prop_initiated vs. tenure*
```{r}
# Smooth the last plot you created of
# of prop_initiated vs tenure colored by
# year_joined.bucket. You can bin together ranges
# of tenure or add a smoother to the plot.

# There won't be a solution image for this exercise.
# You will answer some questions about your plot in
# the next two exercises.

ggplot(data = pf, aes(x = tenure, y = prop_initiated, color = year_joined.bucket)) +
  #geom_line(stat = 'summary', fun.y = median) +
  geom_smooth()
```

*Quiz 8: *
On average, which group initiated the greatest proportion of its Facebook friendships? Use smoother plot from Q7 to answer this.
A: 2012 - 2014
*Quiz 9: *
1. For the group with the largest proportion of friendships initiated, what is the group's average proportion of friendships initiated?
```{r}
summary(subset(pf, year_joined.bucket == '(2012,2014]')$prop_initiated)
```
0.67
2. Why do you think this group's proportion of friendships initiated is higher than the others?
I imagine this group of users can be divided into two general groups: users who joined late compared to their peers, and younger users. For the first group, it makes sense that they have to initiate friendships themselves, as established users might not be aware of them having joined Facebook. This might also to some degree be the case for younger users, as they'll add their older family members. This users however are more likely to have joined around the same time as their peer group, and will likely receive more friend requests than the late joiners.
```{r}
summary(subset(pf, year_joined.bucket == '(2012,2014]'))

by(pf$age,pf$year_joined.bucket, summary)
```

*Quiz 10: Price/Carat Binned, Faceted & Colored*
```{r}
# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

ggplot(data = diamonds, aes(x = cut, y = (price/carat), color = color)) +
  geom_point() +
  facet_wrap(~ clarity) +
  scale_color_brewer(type = 'div') + 
  geom_jitter(width = 0.3, height = 0)

#I like this one better (facet_grid instead of facet_wrap)
ggplot(data = diamonds, aes(x = cut, y = (price/carat), color = color)) +
  geom_point() +
  facet_grid(~ clarity) +
  scale_color_brewer(type = 'div') +
  geom_jitter(width = 0.3, height = 0)

#?geom_jitter()
```

*Quiz 11: *
```{r}
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to continue the investigation you did at the
# end of Problem Set 4 or you can start fresh and choose a different
# data set from Gapminder.

# If you’re feeling adventurous or want to try some data munging see if you can
# find a data set or scrape one from the web.

# In your investigation, examine 3 or more variables and create 2-5 plots that make
# use of the techniques from Lesson 5.

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. the variable(s) you investigated, your observations, and any summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

```

