---
title: "problem_set:_explore_one_variable"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading libraries
library(ggplot2)
```


```{r}
#1a
diamonds
```
```{r}
#1b
ncol(diamonds)

#1c
str(diamonds)

#1d
?diamonds
```

*1. Quiz*
a) There are 53490 observations in the dataset
b) There are 10 variables in the dataset
c) there are 3 ordered factors in the dataset
d) D is the letter for the best diamond color


*2. Quiz: Price Histogram*
# Create a histogram of the price of
# all the diamonds in the diamond data set.

# TYPE YOUR CODE BELOW THE LINE
# =======================================
```{r}
summary(diamonds$price)

qplot(diamonds$price, binwidth = 400) + 
  scale_x_continuous(limits = c(0,20000), breaks = seq(0, 20000, 500))

```
The shape is left skewed. One of the bins, 400 to 800, is more than twice as large as the second largest bin. The mean of the sample is 3933, while the median is 2401. 

*4. Quiz: Diamond Counts*
a) How many diamonds cost less than $300?
```{r}
nrow(subset(diamonds, price < 500))
```
b) How many diamonds cost less than $250?
```{r}
nrow(subset(diamonds, price < 250))
```
c) How many diamonds cost $15000 or more?
```{r}
nrow(subset(diamonds, price >= 15000))

```

```{r}
subset(diamonds$price, diamonds$price > 300)
```

