---
title: "R Notebook"
output: html_notebook
---

*Quiz 1: Price Histograms with Facet and Color*
```{r}
# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

# This assignment is not graded and
# will be marked as correct when you submit.
library(ggplot2)
library(gridExtra)

names(diamonds)

ggplot(data = diamonds, aes(x = price, fill = cut)) + 
  geom_histogram() + 
  facet_wrap(~ color)

#Adjusting for positive skew
ggplot(data = diamonds, aes(x = price, fill = cut)) + 
  geom_histogram() + 
  facet_wrap(~ color) +
  coord_trans(x = "log10") + 
  xlim(100,20000)

```

*Quiz 2: Price vs. Table Colored by Cut*
```{r}
# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

head(diamonds)

ggplot(data = diamonds, aes(x = table, y = price, color = cut)) +
  geom_point() +
  scale_color_brewer(type = 'qual') +
  scale_x_continuous(breaks = seq(50,80,2))

```

*Quiz 3: Typical Table Value*
Use the graph from the last exercise.
What is the typical table range for the majority of diamonds of ideal cut?
53 to 57
What is the typical table range for the majority of diamonds of premium cut?
58 to 62
*Quiz 4: Price vs. Volume and Diamond Clarity*
```{r}
# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')

ggplot(data = subset(diamonds,quantile(x*y*z,0.99) > x*y*z), 
       aes(x = (x*y*z), y = price,  color = clarity)) +
  geom_point() +
  scale_y_log10() +
  scale_color_brewer(type = 'div')

?coord_trans()

```

*Quiz 5: Proportion of Friendships Initiated*
```{r}
# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.
pf <- read.table('pseudo_facebook.tsv', header = 1)
head(pf)

pf <- transform(pf,prop_initiated = friendships_initiated / friend_count)
```

*Quiz 6: prop_initiated vs. tenure*
```{r}
# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg

pf$year_joined <- floor(2014 - (pf$tenure / 365))
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))

ggplot(data = pf, aes(x = tenure, y = prop_initiated, color = year_joined.bucket)) +
  geom_line(stat = 'summary', fun.y = median)
```

*Quiz 7: smoothing prop_initiated vs. tenure*
```{r}
# Smooth the last plot you created of
# of prop_initiated vs tenure colored by
# year_joined.bucket. You can bin together ranges
# of tenure or add a smoother to the plot.

# There won't be a solution image for this exercise.
# You will answer some questions about your plot in
# the next two exercises.

ggplot(data = pf, aes(x = tenure, y = prop_initiated, color = year_joined.bucket)) +
  #geom_line(stat = 'summary', fun.y = median) +
  geom_smooth()
```

*Quiz 8: *
On average, which group initiated the greatest proportion of its Facebook friendships? Use smoother plot from Q7 to answer this.
A: 2012 - 2014
*Quiz 9: *
1. For the group with the largest proportion of friendships initiated, what is the group's average proportion of friendships initiated?
```{r}
summary(subset(pf, year_joined.bucket == '(2012,2014]')$prop_initiated)
```
0.67
2. Why do you think this group's proportion of friendships initiated is higher than the others?
I imagine this group of users can be divided into two general groups: users who joined late compared to their peers, and younger users. For the first group, it makes sense that they have to initiate friendships themselves, as established users might not be aware of them having joined Facebook. This might also to some degree be the case for younger users, as they'll add their older family members. This users however are more likely to have joined around the same time as their peer group, and will likely receive more friend requests than the late joiners.
```{r}
summary(subset(pf, year_joined.bucket == '(2012,2014]'))

by(pf$age,pf$year_joined.bucket, summary)
```

*Quiz 10: Price/Carat Binned, Faceted & Colored*
```{r}
# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

ggplot(data = diamonds, aes(x = cut, y = (price/carat), color = color)) +
  geom_point() +
  facet_wrap(~ clarity) +
  scale_color_brewer(type = 'div') + 
  geom_jitter(width = 0.3, height = 0)

#I like this one better (facet_grid instead of facet_wrap)
ggplot(data = diamonds, aes(x = cut, y = (price/carat), color = color)) +
  geom_point() +
  facet_grid(~ clarity) +
  scale_color_brewer(type = 'div') +
  geom_jitter(width = 0.3, height = 0)


```

*Quiz 11: Gapminder Multivariate Analysis*
```{r}
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to continue the investigation you did at the
# end of Problem Set 4 or you can start fresh and choose a different
# data set from Gapminder.

# If you’re feeling adventurous or want to try some data munging see if you can
# find a data set or scrape one from the web.

# In your investigation, examine 3 or more variables and create 2-5 plots that make
# use of the techniques from Lesson 5.

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. the variable(s) you investigated, your observations, and any summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

library(ggplot2)
#Loading data
fertility <- read.csv('indicator undata total_fertility.csv', header = T, row.names = 1, check.names = F)


head(fertility)

library(reshape2)
#?melt()
fertility.long <- melt(as.matrix(fertility))
names(fertility.long) <- c('country', 'year', 'birth_rate')

#Creating heat map for birth rates, per country and year
ggplot(data = fertility.long, aes(y = country, x = year, fill = birth_rate)) + 
  geom_tile() + 
  scale_fill_gradientn(colors = colorRampPalette(c("blue", "red"))(100)) +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 8)) + 
  ggsave('L8Q11heatmap.png')
```
We see here that there are too many categorical variables for the heatmap to provide much useful information. To improve this I will try to divide the countries into regions and continents instead, using the countrycode library.
```{r}
library(countrycode)
library(dplyr)

names(countrycode_data)

?countrycode

unique(countrycode_data$continent)
unique(countrycode_data$region)

countrycode('Norway', 'country.name', 'region')

fertility$region <- countrycode(row.names(fertility), 'country.name','region')

# # Commenting out some QA which was performed based on countrycode() warning messages
# multimatch_countries = c('Eritrea and Ethiopia','Eastern Africa','Eastern Africa')
# filter(fertility, row.names(fertility) %in% multimatch_countries)
# subset(fertility, row.names(fertility) == 'Channel Islands')
# tail(filter(fertility.long, grepl('ritrea and',country), !is.na(value) ),20)
# filter(fertility, grepl('ritrea and',row.names(fertility)) )#, !is.na(value) )
# filter(countrycode_data, grepl('ersey',country.name.en))
fertility$continent <- countrycode(row.names(fertility), 'country.name','continent')

#Splitting continent value 'Americas' into North and South
fertility$continent[fertility$region == 'South America'] <- 'South America'
fertility$continent[fertility$continent == 'Americas'] <- 'North America'

count(fertility, continent)

#row.names(subset(fertility, is.na(continent)))
#row.names(subset(fertility, is.na(region)))

#Adding region and continent to fertility.long
library(tibble)
fertility.long <- melt(rownames_to_column(fertility,'country'), 
                       id = c('country','region','continent'))
names(fertility.long)[4:5] <- c('year','birth_rate')

ggplot(fertility.long, aes(x = region) )+ 
  geom_histogram(stat = 'count')

ggplot(fertility.long, aes(x = continent) )+ 
  geom_histogram(stat = 'count')

filter(fertility.long, region == 'Western Europe' & is.na(birth_rate))

ggplot(data = subset(fertility.long, !is.na(region)), 
       aes(x = year, y = region, fill = birth_rate)) + 
  geom_tile() + 
  scale_fill_gradientn(colors = colorRampPalette(c("blue", "red"))(100)) +
      theme(axis.text.x = element_text(size = 10),
          axis.text.y = element_text(size = 10)) + 
  ggsave('L8Q11heatmapregion.png')

ggplot(data = subset(fertility.long, !is.na(continent)), 
       aes(x = year, y = continent, fill = birth_rate)) + 
  geom_tile(na.rm = TRUE) + 
  scale_fill_gradientn(colors = colorRampPalette(c("blue", "red"))(100)) +
      theme(axis.text.x = element_text(size = 10),
          axis.text.y = element_text(size = 10)) + 
  ggsave('L8Q11heatmapcontinent.png')

by(fertility.long$birth_rate, fertility.long$region, summary)

?geom_tile()

subset(fertility.long, region == 'Southern Europe')

```

This also does not tell us much, and I suspect the tile plot is not very good to use for this dataset. Next I will do some wrangling to get a better look at how the mean birth rate for the different continents varied over time.
```{r}
#Adding bins for years

head(fertility.long2)

#First plotting average per year to see if I can make some bins wider than others
p1 <- ggplot(data = subset(fertility.long, !is.na(birth_rate) ),
                     aes(y = birth_rate, x = as.factor(year) ) ) +
  geom_boxplot()

p2 <- ggplot(data = subset(fertility.long, !is.na(birth_rate)),
                     aes(y = birth_rate)) +
  geom_line(aes(x = as.numeric(as.character(year)) ),
             stat = 'summary', fun.y = mean, color = 'blue') +
 geom_line(aes(x = as.numeric(as.character(year)) ),
                stat = 'summary', fun.y = median, color = 'red')

grid.arrange(p1,p2, ncol = 1)
# ggsave('L8Q11grid.png', arrangeGrob(p1,p2, ncol = 1)) #to save grid plot

#Changing type of year for cutting purposes
fertility.long$year <- as.numeric(as.character(fertility.long$year))

#?cut()
#Based on the above, I'll make the pre-1900 bins/buckets fairly wide
fertility.long$year.bucket <- cut(fertility.long$year,breaks = c(
  1800,1875,seq(1900,2020,10) ), dig.lab = 4, right = FALSE)

ggplot(data = subset(fertility.long, !is.na(birth_rate) & !is.na(continent) ), 
       aes(x = year.bucket, y = birth_rate)) + 
  geom_bar(stat = 'summary', fun.y = mean) +
  facet_wrap(~ continent) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9)) +
  scale_y_continuous(sec.axis = dup_axis()) #+ 
#  ggsave('L8Q11facet.png')
```
With the exception of Europe, the mean birth rates for the different continents were fairly stable at around 6 until 1970. Europe's mean birth rate was already as low as 5 by the end of the 1800s, and steadily declined to a below 3 by the 1970s.The decline slowed down a bit after that, to a low point of around 1.75 in the first decade of the 2000s, and actually went up a little bit in the first half of the 2010s. For Asia, Oceania, North American and South America, a rapid decline in mean birth rate started in the 1970s, and continued into the first halft of the 2010s, when it was between 2 and 2.6 for the four continents. Africa has also had a a decline in mean birth rate, but it did not really start until the 1990s, and by the first half of the 2010s the mean birth rate for African countries was still above 4. 

A flaw in the plots above and the following analysis is that it is not weighted for country population, leading to a tiny country having the same impact as the most populous countries. I started trying to match population data from the wpp2015 package (see below) to the country names from the fertility data set, and ran into a severe matching issue. 85 of the 258 country names in the fertility dataset did not have a match on name in the wpp2015 dataset. Sometime this was due to the countries being very small (e.g. Andorra), but often this was due to different naming conventions in the different datasets (e.g. Russia and Russian Federation). In the end I decided this to be outside the scope of this exercise, and decided to stop at this point.

```{r}
#install.packages('wpp2015')
library(wpp2015)

data(pop)
head(pop)

unique(anti_join(fertility.long, pop, by = c('country' = 'name'))$country)
length(unique(anti_join(fertility.long, pop, by = c('country' = 'name'))$country))
unique(anti_join(pop, fertility.long, by = c('name' = 'country'))$name)
length(unique(fertility.long$country))

subset(pop, name %in% 'Russian Federation')
subset(pop, name == 'Andorra')
arrange(pop,name)$name
```
