---
title: "problem_set:_explore_one_variable"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading libraries
library(ggplot2)
```


```{r}
#1a
diamonds
```
```{r}
#1b
ncol(diamonds)

#1c
str(diamonds)

#1d
?diamonds
```

*1. Quiz*
a) There are 53490 observations in the dataset
b) There are 10 variables in the dataset
c) there are 3 ordered factors in the dataset
d) D is the letter for the best diamond color


*2. Quiz: Price Histogram*
# Create a histogram of the price of
# all the diamonds in the diamond data set.

# TYPE YOUR CODE BELOW THE LINE
# =======================================
```{r}
summary(diamonds$price)

qplot(diamonds$price, binwidth = 400) + 
  scale_x_continuous(limits = c(0,20000), breaks = seq(0, 20000, 500))

```
The shape is left skewed. One of the bins, 400 to 800, is more than twice as large as the second largest bin. The mean of the sample is 3933, while the median is 2401. 

*4. Quiz: Diamond Counts*
a) How many diamonds cost less than $300?
```{r}
nrow(subset(diamonds, price < 500))
```
b) How many diamonds cost less than $250?
```{r}
nrow(subset(diamonds, price < 250))
```
c) How many diamonds cost $15000 or more?
```{r}
nrow(subset(diamonds, price >= 15000))

```

```{r}
subset(diamonds$price, diamonds$price > 300)
```

*5. Quiz: Cheaper Diamonds*
# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won’t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

```{r}
qplot(diamonds$price, binwidth = 25, color = I('black'), fill = I('#099DD9') ) + 
  scale_x_continuous(limits = c(350,1200), breaks = seq(300, 1200, 50))

#Rewriting the same in ggplot
ggplot(aes(x = price), data = diamonds) + 
  geom_histogram(binwidth = 25, color = I('black'), fill =  '#F79420', boundary = 1 ) +
  scale_x_continuous(limits = c(350,1201), breaks = seq(350, 1200, 50)) +
  xlab("Price (dollars)") + ylab("Number of diamonds")

ggsave('cheaper_diamonds_350-1200.png')

ggplot(aes(x = price), data = diamonds) + 
  geom_histogram(binwidth = 1, color = I('black'), fill =  '#F79420' ) +
  scale_x_continuous(limits = c(700,776), breaks = seq(700, 775, 5)) +
  xlab("Price (dollars)") + ylab("Number of diamonds")

ggsave('cheaper_diamonds_700-775.png')

#qplot(c(18,11,11,11,12,12,13,14,15,19,19,19,19,20,21,21), binwidth=1, color=I('black'), boundary = 1) + 
# scale_x_continuous(breaks = seq(10,20,2), limits = c(10,20))

nrow(subset(diamonds,price >= 350 & price < 375))
nrow(subset(diamonds,price > 1200 & price < 1225))
```

*6. Quiz: Price by  Cut Histograms*
# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.

```{r}
ggplot(aes(x = price), data = diamonds) + 
  geom_histogram(binwidth = 10) + 
  scale_x_continuous() + #limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  facet_wrap(~cut, ncol = 3)
```
The distributions are similar, except for 'Fair' cut. That doesn't tell us much, as the size of the stones are not taken into account.

*7. Quiz: Price by Cut*
a) Which cut has the highest priced diamond?
```{r}
#method 1: If few enough categories, just look at the summary output
by(diamonds$price, diamonds$cut, summary)
#NOTE: max is rounded

#method 2:
head(diamonds[with(diamonds,order(-price)),])

#method 3:
diamonds[which.max(diamonds$price),]
```
Premium cut has the highest priced diamond with a price of $18823.

b) Which cut has the lowest priced diamond?
```{r}
head(diamonds[with(diamonds,order(price)),])

#Alternative method (will not find both)
diamonds[which.min(diamonds$price),]
```
Ideal and Premium cuts both have a diamond priced at $326, which is the lowest price point.

c) Which cut has the lowest median price?
```{r}
by(diamonds$price, diamonds$cut, median)
```
Ideal has the lowest median price.

BONUS QUESTION (I answered this before C due to incorrect transcription)
d) Which cut has the median price?
```{r}
#Simplest solution, requires more steps
subset(diamonds, price == median(diamonds$price))

#More complex solution(?)
diamonds[which(diamonds$price == median(diamonds$price)),]
#To get single row
diamonds[which.min(abs(diamonds$price - median(diamonds$price))),]
#Get all rows, but just the unique cut values
unique(diamonds[which(diamonds$price == median(diamonds$price)),2])
```
All the categories: Premium, Ideal, Very Good, Good and Fair

*8. Quiz: Scales and multiple histograms*
# In the two last exercises, we looked at
# the distribution for diamonds by cut.

# Run the code below in R Studio to generate
# the histogram as a reminder.

# ===============================================================
```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut)
```
# ===============================================================

# In the last exercise, we looked at the summary statistics
# for diamond price by cut. If we look at the output table, the
# the median and quartiles are reasonably close to each other.

# diamonds$cut: Fair
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     337    2050    3282    4359    5206   18570 
# ------------------------------------------------------------------------ 
# diamonds$cut: Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     327    1145    3050    3929    5028   18790 
# ------------------------------------------------------------------------ 
# diamonds$cut: Very Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     336     912    2648    3982    5373   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Premium
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326    1046    3185    4584    6296   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Ideal
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326     878    1810    3458    4678   18810 

# This means the distributions should be somewhat similar,
# but the histograms we created don't show that.

# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.
```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = 'free_y')
#Same in ggplot
ggplot(aes(price), data = diamonds) + 
  geom_histogram() + 
  facet_wrap(~cut, ncol = 2, scales = 'free_y')
```

*Quiz 9: Price per carat by cut*
# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

# Submit your final code when you are ready.
```{r}
#first attempt
ggplot(aes(log10(price/carat)), data = diamonds) + 
  geom_histogram(binwidth = 0.1) + 
  facet_wrap(~cut)

#more proper way
ggplot(aes(price/carat), data = diamonds) + 
  geom_histogram(binwidth = 0.05, color = 'black') + 
  scale_x_log10() + 
  facet_wrap(~cut) + 
  xlab("Price, logarithmic scale")

```
*10. Quiz: Price Box Plots*
# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won’t be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# =================================================================
```{r}
ggplot(aes(x = color, y = price), data = diamonds) +
  geom_boxplot(fill = c('gray100','gray99','gray98','gray97','gray96','gray95','floralwhite')) + 
  #  geom_boxplot(fill = c('azure','azure1','ghostwhite','floralwhite','floralwhite','cornsilk','cornsilk')) + 
  coord_cartesian(ylim = c(0,8000)) + 
  scale_y_continuous(breaks = seq(0,8000,500))

ggsave('boxplot_diamond_color.png')

by(diamonds$price, diamonds$color, summary)

```
Purely based on color, the diamond prices seems to fluctuate more for the lower color grade than for the higher grades.

*Quiz 11: Interquartile range - IQR*
```{r}
#IQR for diamonds with color D (best color)
IQR(x = subset(diamonds, color == 'D')$price)
4214 - 911

#IQR for diamonds with color J (best color)
IQR(x = subset(diamonds, color == 'J')$price)
7695 - 1860
```

*12. Quiz: Price per Carat Box Plots by Color*
# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.

# Go to the discussions to
# share your thoughts and to discover
# what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.

# SUBMIT YOUR CODE BELOW THIS LINE
# ===================================================================
```{r}
ggplot(aes(x = color, y = price/carat), data = diamonds) +
  geom_boxplot(fill = c('gray100','gray99','gray98','gray97','gray96','gray95','floralwhite')) + 
  geom_boxplot(fill = c('azure','azure1','ghostwhite','floralwhite','floralwhite','cornsilk','cornsilk')) + 
  coord_cartesian(ylim = c(0,6000)) + 
  scale_y_continuous(breaks = seq(0,6000,500))

by(diamonds$price/diamonds$carat, diamonds$color, summary)

#Best plots from forum
uniform<-subset(diamonds, cut=="Ideal" & clarity == "IF" & carat > 0.7)
qplot(y=price/carat, x=color,  data=uniform, geom="boxplot", fill=color)
uniform<-subset(diamonds, cut=="Ideal" & clarity == "IF")

table(uniform$color)
qplot(x = color, y = price/carat, 
data = uniform,
geom = "boxplot",
fill = color)

#from forum
#The code/plot below shows the plot from the original post with a variable box width:
qplot(x = color, y = price/carat, 
data = uniform,
geom = "boxplot", 
varwidth = TRUE, # added this
fill = color)

qplot(data = diamonds, x = clarity, y = price/carat,
xlab = 'Clarity',
ylab = 'Price',
geom = 'boxplot',
fill=color) +
coord_cartesian(ylim=c(350,18700))
```

```{r}
ggplot(aes(x = color, y = carat), data = diamonds) +
  geom_boxplot(fill = c('gray100','gray99','gray98','gray97','gray96','gray95','floralwhite')) + 
  geom_boxplot(fill = c('azure','azure1','ghostwhite','floralwhite','floralwhite','cornsilk','cornsilk'))
 # coord_cartesian(ylim = c(0,6000)) + 
#  scale_y_continuous(breaks = seq(0,6000,500))
```

*Quiz 13: Carat Frequency Polygon*
```{r}
summary(diamonds$carat)

table(diamonds$carat)

ggplot(aes(x = carat), data = diamonds) + 
  geom_freqpoly(binwidth = 0.1) + 
  scale_x_continuous(breaks = seq(0,5,0.1)) +
  scale_y_continuous(breaks = seq(0,15000,1000))

ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')

```
Question focuses on binwidth. 0.3, 0.8 and 1.01 has a count above 2000, the rest of the options do not.

=========================================
*Data Wrangling with R*
Full preso at https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf
```{r}
library(dplyr)
#install.packages('devtools')
#devtools::install_github('rstudio/EDAWR')
library(EDAWR)

#These do the same thing
select(tb, child:elderly)
tb %>% select(child:elderly)

storms
storms$storm

names(cases)[-1]
unlist(cases[1:3, 2:4])

storms$pressure / storms$wind

```
Tidy data
1. Each variable is stored in its own column.
2. Each observation is saved in its own row.
3. Each "type" of observation stored in a single table (here, storms).

tidyr
-A package that reshapes the layout of tables
-Two main functions: gather() and spread()

```{r}
#install.packages("tidyr")
library(tidyr)
?gather
?spread
```

Gather
Collapses multiple columns into two columns:
1. a *key* column that contains the former column names
2. a *value* column that contains the former column cells
```{r}
cases
gather(cases, "year","n", 2:4)
```

Spread
1. Each unique value in the *key* column becomes a column name
2. each value in the *value* column becomes a cell in the new columns
```{r}
pollution
spread(pollution, size, amount)
```

unite() and separate()
```{r}
?unite
?separate
```

Separate
Separate splits a column by a character string separator.
```{r}
storms
storms2 <- separate(storms, date, c("year", "month", "day"), sep = "-")
storms2
```

Unite
Unite unites columns into a single column.
```{r}
unite(storms2, date, year, month, day, sep = "-")
```

dplyr: A package that helps transform tabular data.

```{r}
library(dplyr)
?select()  # Extract existing variables
?filter()  # Extract existing observations
?arrange  
?mutate    # Derive new variables (from existing variables)
?summarise # Change the unit of analysis
?group_by
```

nycflights13
Data sets related to flights that departed from NYC in 2013
```{r}
#install.packages("nycflights13")
library(nycflights13)

?airlines
?airports
?flights
?planes
?weather
```

Select
Useful select functions:
contains(), ends_with(), everything(), matches(), 
num_range(), one_of(), starts_with()
```{r}
#similar to SQL: place 1 is table, rest is columns
select(storms, storm, pressure)
#all but
select(storms, -storm)
#from column to column
select(storms, wind:date)

```

Filter
```{r}
#Select with WERE clause
filter(storms, wind >= 50)
filter(storms, wind >= 50, 
       storm %in% c("Alberto", "Alex", "Allison"))
```

Mutate
```{r}
#Adds new column
mutate(storms, ratio = pressure / wind)
mutate(storms, ratio = pressure / wind, inverse = ratio ^-1)
```

Summarise
```{r}
pollution %>% summarise(median = median(amount), variance = var(amount))
pollution %>% summarise(mean = mean(amount), sum = sum(amount), n = n())
```

Arrange
```{r}
#Sorts by wind column
arrange(storms, wind)
arrange(storms, desc(wind))
#Sorts by wind, then by date
arrange(storms, wind, date)
```

Useful use of pipe operator %>%
Shortcut on mac: cmd + shift + M
```{r}
storms %>% filter(wind >= 50) %>% select(storm, pressure)
```

*Unit of Analysis*
```{r}
pollution %>% group_by(city)
pollution %>% group_by(city) %>% 
  summarise(mean = mean(amount), sum = sum(amount), n = n())

#by size
pollution %>% group_by(size) %>% summarise(mean = mean(amount))

#Ungroup
pollution %>% ungroup()

tb %>% group_by(country, year) %>% 
  summarise(cases = sum(child))

tb %>% group_by(country, year) %>% 
  summarise(cases = sum(child)) %>%
  summarise(cases = sum(cases))

```

*Joining Data*

dplyr::bind_cols()
```{r}
y
z
bind_cols(y, z)
```

dplyr::bind_rows()
```{r}
bind_rows(y, z)
```

dplyr::union()
```{r}
union(y, z)
```

dplyr::intersect()
```{r}
intersect(y, z)
```

dplyr::setdiff()
```{r}
setdiff(y, z)
```

dplyr::left_join()
```{r}
songs
artists
left_join(songs, artists, by = 'name')

left_join(songs2, artists2, by = c("first", "last"))
```

dplyr::inner_join()
```{r}
inner_join(songs, artists, by = 'name')
```

dplyr::semi_join()
```{r}
#Only returns columns from left table, matches same as inner join
semi_join(songs, artists, by = 'name')
```

dplyr::anti_join()
```{r}
#left join where b.key IS NULL
anti_join(songs, artists, by = 'name')
```
=========================================

*15. Quiz: Gapminder Data*
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.

# You might use a simple histogram, a boxplot split over a categorical variable,
# or a frequency polygon. The choice is yours!

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# ====================================================================================

```{r}
#setwd("/Users/ggaasemyr/Documents/udacity/ipython_notebooks/p4/data_analysis_with_R/")

fertility <- read.csv('indicator undata total_fertility.csv', header = T, row.names = 1, check.names = F)


pollution

ggplot(aes(x = city, y = pollution$'amount'), data = pollution) + 
  geom_point()

ggplot(aes(x = rownames(fertility), y = fertility$'2015'), data = fertility) + 
  geo()

fertility$'2015'

qplot(fertility$'2015', data = fertility)

summary(fertility$'2015')

View(fertility[-3])


fertility_mean_by_year = colMeans(fertility, na.rm = T)

ggplot(aes(x = ind, y = values), data = stack(fertility)) + 
  geom_boxplot() + 
  scale_y_continuous()

ggplot(aes(x = values), data = stack(fertility)) + 
  geom_bar()

ggplot(aes(x = city), data = pollution) + 
  geom_bar()

ggplot(aes(x))

stack((fertility_mean_by_year))

ggplot(aes(x = ind, y = values), data = stack((fertility_mean_by_year))) + 
  geom_line()

?stack(fertility)

ggplot(aes(x = ind, y = values, group = 1, color = I('blue')), data = stack(fertility_mean_by_year)) +
  geom_line() 

ggplot(aes(x = ind, y = values, color = I('black')), fill = rgb(r,g,b), data = stack(fertility)) +
  geom_boxplot() 

fertility[200:215]

#CONTINUE ON THIS
summary(fertility$'2015')[5] - summary(fertility$'2015')[2]

summary(fertility$'2015')[5,]
summary(fertility$'2015')[2,]

3.495-1.785

ggplot(stack(fertility[200:215]), aes(x = ind, y = values)) + 
  geom_boxplot()


ggplot(fertility, aes(x = region, y = fertility$'2015')) + 
  geom_boxplot()

ggplot(fertility, aes(x = continent, y = fertility$'2015')) + 
  geom_boxplot()

install.packages("countrycode")
library(countrycode)

countrycode(c('norway','usa'),'country.name','region')

countrycode(row.names(fertility),'country.name','region')

?countrycode_data()

fertility$continent <- countrycode(row.names(fertility),'country.name','continent')
fertility$region <- countrycode(row.names(fertility),'country.name','region')

fertility[(length(fertility)-2):length(fertility)]

View(tail(countrycode_data))

fertility[2:3]

list(row.names(fertility))# typeof(fertility)

View(filter(countrycode_data,cowc == 'USA'))

distinct(countrycode_data,eurocontrol_pru)
distinct(countrycode_data,region)
distinct(countrycode_data,continent)

```

16. Quiz: Exploring Your Friends' Birthdays

# Your task is to investigate the distribution of your friends'
# birth months and days.

# Here some questions you could answer, and we hope you think of others.

# **********************************************************************

# How many people share your birthday? Do you know them?
# (Reserve time with them or save money to buy them a gift!)

# Which month contains the most number of birthdays?

# How many birthdays are in each month?

# Which day of the year has the most number of birthdays?

# Do you have at least 365 friends that have birthdays on everyday
# of the year?

# **********************************************************************

# You will need to do some data munging and additional research to
# complete this task. This task won't be easy, and you may encounter some
# unexpected challenges along the way. We hope you learn a lot from it though.

# You can expect to spend 30 min or more on this task depending if you
# use the provided data or obtain your personal data. We also encourage you
# to use the lubridate package for working with dates. Read over the documentation
# in RStudio and search for examples online if you need help.

# You'll need to export your Facebooks friends' birthdays to a csv file.
# You may need to create a calendar of your Facebook friends’ birthdays
# in a program like Outlook or Gmail and then export the calendar as a
# csv file.

# Once you load the data into R Studio, you can use the strptime() function
# to extract the birth months and birth days. We recommend looking up the
# documentation for the function and finding examples online.

# We've included some links in the Instructor Notes to help get you started.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation below the line. Submit it when you are ready.
# ===============================================================================

```{r}
friend_birthdays <- read.csv("fb_friend_birthdays.csv")

friend_birthdays$date = strptime(friend_birthdays$Start, format = "%m/%d/%y %H:%M")

qplot(friend_birthdays$date, bins=365)
```

