TITLE by YOUR_NAME_HERE
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
#install.packages('ggplot2')
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data

setwd('/Users/ggaasemyr/Documents/udacity/ipython_notebooks/p4/')

characters <- read.csv('data/character_list5.csv', strip.white = TRUE)

?read.csv()

?read.csv

head(characters)
str(characters)
movies <- read.csv('data/meta_data7.csv', encoding = 'UTF-8')
head(movies)
str(movies)

View(movies)

```

Joining character data with movies, and summarizing word count per movie
```{r}
#ensuring all movies have a match
anti_join(characters, movies, by='script_id')

nrow(characters)
#Adding movie data to characters
characters <- inner_join(characters, movies, by='script_id')
#ensuring same count
nrow(characters)

head(characters)

#Summarizing word counts per movie
script_summary <-
  group_by(characters, script_id, title) %>% 
  summarise(total_words = sum(words), characters = n()) %>% 
  arrange(desc(total_words))

```

Loading academy awards data
```{r}
academy.nominated <- read.csv('data/database.csv', strip.white = TRUE)
str(academy.nominated)
academy.nominated$Year_numeric <- as.numeric(as.character(academy.nominated$Year))
academy.winners <- subset(academy.nominated, Winner == 1)
```


```{r}
View(characters)

percentile <- ecdf(as.numeric(script_summary$total_words))

percentile(filter(script_summary, title == 'Before Sunset')$total_words)

nrow(movies)

arrange(characters, desc(words))

filter(master.list, grepl('inklater', master.list$director)) %>% 
  inner_join(script_summary, by = 'title')

#Checking for alternative title matches
View(inner_join(filter(master.list, as.character(title) != 
         as.character(alternate_title)), movies, 
                by = c('alternate_title' = 'title')) )

```

### Dataset Limitations
After having worked with the datasets for a while, I have discovered that the 
datasets provided by Pudding have some limitations. I will highlight some of 
them in the next section

> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

I want to take a look at how lines in movie scripts affect chances of winning 
academy awards (AKA Oscar) and also how lines affect IMDB scores. Ideally I 
would like to evaluate the vocabulary in scripts for oscar winning movies, 
compared to non-winning movies. Obtaining such has however turned out to be 
very difficult, so I have instead decided to use datasets containing 
word and character counts.

```{r}
View(academy.winners %>%  
  group_by(Award) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)))

str(academy.winners)

View(filter(academy.winners, Year == 2014))

tail(academy.winners)


filter(anti_join(script_summary, movies, by = 'script_id',
           suffix = c('','.y'), 'name' == 'name.y')))

script_summary <- inner_join(script_summary, movies, by = 'script_id',
           suffix = c('','.y'))[c('script_id','imdb_id','title', 'total_words',
                                  'characters','year')]

View(script_summary)

str(movies)

head(filter(academy.winners, Year_numeric >= 1995, 
         Year_numeric <= 2015))

academy.winners.1995.to.2015 <- 
  filter(academy.winners, Year_numeric >= 1995, 
         Year_numeric <= 2015)

```

```{r}
#Joining all categories

awards.2015 <- (filter(academy.winners, Year == '2015')$Award)

View(academy.winners.1995.to.2015)

awards.count <- 
  academy.winners.1995.to.2015 %>% 
  group_by(Award) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

awards.count$in_2015 <- awards.count$Award %in% awards.2015
View(awards.count)
View(filter(awards.count, in_2015 == TRUE))

View(awards)

```
Looking at the aggregated awards, it becomes clear that some have 
changed names in recent years, and are for all purposes the same. 
Therefore I will rename some of the older award names in the 
academy.winners.1995.to.2015 dataframe to the new award names, 
in the following way:
- 'Writing (Screenplay Based on Material Previously Produced or Published)' 
will be replaced by 'Writing (Adapted Screenplay)'
- 'Writing (Screenplay Written Directly for the Screen)' will
be replaced by 'Writing (Original Screenplay)'
-'Music (Original Dramatic Score)' and 'Music (Original Musical 
or Comedy Score)' will be replaced by 'Music (Original Score)'
-'Sound' will be replaced by 'Sound Mixing'
-'Makeup' will be replaced by 'Makeup and Hair Styling'
```{r}
#Creating dictionary for award name replacements
award_replacements <- c(
  'Writing (Screenplay Based on Material Previously Produced or Published)'='Writing (Adapted Screenplay)',
  'Writing (Screenplay Written Directly for the Screen)'=
    'Writing (Original Screenplay)',
 'Music (Original Dramatic Score)'='Music (Original Score)',
 'Music (Original Musical or Comedy Score)'='Music (Original Score)',
 'Sound'='Sound Mixing',
 'Makeup'='Makeup and Hairstyling')

count = 0
for (name in names(award_replacements)) {
  count = count + 1

  #Test
  print(name)
  print(nrow(academy.winners.1995.to.2015[
    (academy.winners.1995.to.2015$Award == name),]))

  #Changing name to 2015 names
  academy.winners.1995.to.2015[(academy.winners.1995.to.2015$Award == 
                                 name),]$Award <- award_replacements[count]
}

```
Now that the Award names are updated in the award winner 1995-2015 dataframe, 
I will create updated summary tables as well.
```{r}
awards.edited.count <- 
  academy.winners.1995.to.2015 %>% 
  group_by(Award) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

awards.edited.count$in_2015 <- awards.edited.count$Award %in% awards.2015
View(awards.edited.count)
View(filter(awards.edited.count, in_2015 == TRUE))

nrow(filter(awards.edited.count, count > 20))
nrow(filter(awards.count, count > 20))
```
We now have 21 award categories for the 21 year period 1995 to 2015, compared 
to 16 before we updated the award names to the 2015 Academy award ceremony.

Next I will join the academy award data with the script data set, and see how
many years we have script data for for the different award categories. 
```{r}
#Joining on name
combined.1995.to.2015 <- 
   academy.winners.1995.to.2015 %>% 
   inner_join(filter(script_summary,
                     year >= 1995, year <=2015), 
              by = c('Name' = 'title'))

#Adding rows joining on film
combined.1995.to.2015 <- academy.winners.1995.to.2015 %>% 
   inner_join(filter(script_summary,
                     year >= 1995, year <=2015), 
              by = c('Film' = 'title')) %>% 
  union(combined.1995.to.2015)

#Checking for errors
View(filter(combined.1995.to.2015, year != Year)[,c(1,12,3,5),])

#Removing error match
filter(combined.1995.to.2015, Name == 'Feast')
combined.1995.to.2015 <- filter(combined.1995.to.2015, Name != 'Feast')
#Ensuring correct count
nrow(combined.1995.to.2015)
```

```{r}
#Finding award categories with no matching script data
no.award.matches <- academy.winners.1995.to.2015 %>% 
  anti_join(combined.1995.to.2015, by='Award') %>%  
  group_by(Award) %>% 
  summarise(n = n())
#Taking a look at the missing categories
no.award.matches

#Potential awards to include
join_awards <- filter(awards.edited.count, count > 20 & count < 30 & 
                        !(Award %in% no.award.matches$Award))$Award

#Looking for suspicious missing Oscar winners
View(
  academy.winners.1995.to.2015 %>% 
    filter(Award %in% join_awards & Award != 'Foreign Language Film') %>% 
    anti_join(combined.1995.to.2015, by = c('Award', 'Year')) %>% 
    arrange(desc(Year,Name)))

```
Manual review of the filtered list above revealed a few more matches:
Precious - Precious: Based on the Novel 'Push' by Sapphire
The Chronicles of Narnia: The Lion, the Witch and the Wardrobe - 
The Chronicles of Narnia: The Lion, the Witch and the Wardrobe (trailing space)
Adaptation. - Adaptation
```{r}

script_replace_academy_name <- c(
"Precious: Based on the Novel 'Push' by Sapphire" = 'Precious',
'The Chronicles of Narnia: The Lion, the Witch and the Wardrobe ' =
'The Chronicles of Narnia: The Lion, the Witch and the Wardrobe')

academy_replace_script_name <- c('Adaptation' = 'Adaptation.')

filter(academy.winners, grepl('Narnia', academy.winners$Name))
filter(academy.winners, Name == 'The Chronicles of Narnia: The Lion, the Witch and the Wardrobe ')
filter(academy.winners, grepl('Narnia', academy.winners$Film))
filter(combined.1995.to.2015, Film == 'The Chronicles of Narnia: The Lion, the Witch and the Wardrobe')
filter(script_summary, title == 'The Chronicles of Narnia: The Lion, the Witch and the Wardrobe') %>% 
  inner_join(academy.winners, by = c('title' = 'Name'))

```

Checking for trailing spaces.
```{r}
#install.packages('stringr')
library(stringr)

removed_space_trail <- filter(academy.winners.1995.to.2015, str_sub(Name, -1) == " ")$Name %>% 
  str_sub(,-2)

removed_space_trail

filter(script_summary, title %in% removed_space_trail)

#Also checking for movies ending with a period (.)
removed_period_trail <- filter(academy.winners.1995.to.2015, str_sub(Film, -1) == ".")$Film %>% 
  str_sub(,-2) %>% 
  union(
    removed_period_trail <- filter(academy.winners.1995.to.2015, str_sub(Name, -1) == ".")$Name %>% 
  str_sub(,-2)
  )

removed_period_trail
```
Only the Narnia movie gave a match, which I already had on my list of movie 
names to change. No need to make further changes based on trailing spaces then. 
Only actor names and a song were returned for the period trail, so no need to 
check for movie matches.

```{r}


#Changing the names
academy.winners.1995.to.2015$Film <- recode(academy.winners.1995.to.2015$Film, !!!as.list(script_replace_academy_name))
academy.winners.1995.to.2015$Name <- recode(academy.winners.1995.to.2015$Name, !!!as.list(script_replace_academy_name))

tmp_more_matches <- 
  academy.winners.1995.to.2015 %>% 
  inner_join(filter(script_summary,
                     year >= 1995, year <=2015), 
              by = c('Name' = 'title')) %>% 
  rbind(
    academy.winners.1995.to.2015 %>% 
    inner_join(filter(script_summary,
                     year >= 1995, year <=2015), 
              by = c('Film' = 'title')))

combined.1995.to.2015  <- 
  rbind(combined.1995.to.2015,
    anti_join(filter(tmp_more_matches, Name != 'Feast'),
                  combined.1995.to.2015, by = c('Award', 'Year'))
        )

combined.1995.to.2015[movies$title %in% combined.1995.to.2015$Film]
#[add some more arranging above, and merge film and name]
```


It is now time to check how many Academy award winning movies from 1995 to 2015 
which exist in the script dataset. I will probably set a threshold between 10 
and 15 winners for including an award category in the final dataset.
```{r}
combined.count <- 
  combined.1995.to.2015 %>% 
  group_by(Award) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

combined.count$in_2015 <- combined.count$Award %in% awards.2015
View(combined.count)
View(filter(combined.count, in_2015 == TRUE))
```


```{r}
View(filter(academy.winners, Ceremony == 88))#Winner == 'The Revenant')

```

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}

#plotting correlation between character count and total words
ggplot(data = script_summary, aes(x = characters, y = total_words)) +
  geom_point() + 
  geom_smooth() + 
  geom_line(stat = 'summary', fun.y = mean, color = 'red')

subset(script_summary, title == 'Cast Away')
subset(characters, title == 'Cast Away')

?cor()

cor(script_summary$total_words, script_summary$characters, method = 'pearson')
```

[MOVIES BY YEAR PLOT]
```{r}
ggplot(movies, aes(x = year)) + 
  geom_histogram(bins = length(seq(min(movies$year), 
                                   max(movies$year), 1)), 
                 color = 'red', alpha = .2) +
  scale_x_continuous(breaks = seq(1920, 2015, 5))

```


```{r}
# academy.nominated <- read.csv('data/database.csv', strip.white = TRUE)
# str(academy.nominated)
# academy.nominated$Year_numeric <- as.numeric(as.character(academy.nominated$Year))
# academy.winners <- subset(academy.nominated, Winner == 1)


head(semi_join(by = c('title' = 'Film'), #| c('title' = 'Name'),
                      script_summary,academy.winners))

tail(inner_join(by = c('title' = 'Film'), #| c('title' = 'Name'), 
           script_summary,academy.nominated)  %>% 
  inner_join(by = c('title' = 'Name'), academy.nominated), 20)

filter(academy.nominated, Film == 'Tootsie')

inner_join(by = c('title' = 'Film'), #| c('title' = 'Name'), 
           script_summary,academy.nominated) %>% 
  inner_join(by = c('title' = 'Name'), academy.nominated) %>% 
  group_by(title) %>% 
  summarise(year = first(Year), nominations = n(), oscars = n_distinct(Winner)) %>% 
  arrange(desc(year)) %>% 
  head(20)

academy.nominated %>% 
  group_by(Film)


academy.nominated %>% 
  group_by(Film) %>% 
  summarise(nominations = n() )#, wins = n_distinct(Winner)) %>% 
  arrange(desc(wins))

filter(academy.nominated, Name == 'Titanic' | Film == 'Titanic')

# filter(academy.nominated, grepl('Titanic ',Name))

#checking for strange misses
View(anti_join(academy.winners, script_summary, by = c('Film' = 'title') ) %>% 
  arrange(desc(Year)))

repl
?sub

ifelse()

View(tail(academy.winners,20))

head(academy.nominated[,1:5])
head(academy.nominated[,c(1:4,6)])

?union(academy.nominated[,1:5], academy.nominated[,c(1:4,6)])

str(academy.nominated$Award)
df %>%
  group_by(name) %>%
  filter(n() < 4)

group
group_by(academy.nominated, Award) %>% 
  filter(n() < 10) %>% 
  group_by(Year) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))

ggplot(group_by(academy.nominated, Award) %>% filter(n() < 10),
       aes(x = Award) ) +
  geom_bar(stat = 'count') +
  geom_text(stat='count',aes(label=..count..),vjust=-1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 9))

?geom_text
```

[WIKIPEDIA DATASET]
```{r}
all.oscar.winners <- read.csv2('data/all_oscar_winners.tsv', sep = '\t')

head(all.oscar.winners[,1:2])
#Count of oscar winners
nrow(all.oscar.winners)

#Matching Oscar winners with Pudding dataset
oscar_matches <- inner_join(
  filter(all.oscar.winners, Year < 2016), 
  movies, by = c('Film' = 'title'))

#Removing matches with years far apart
oscar_matches <- oscar_matches %>% 
  filter(as.numeric(Year) <= as.numeric(year) + 2 & 
                 as.numeric(Year) >= as.numeric(year) - 2)
#Count of matches
nrow(oscar_matches)

#Missing pre-2016 oscar winners in Pudding dataset
missing.oscar.scripts <- anti_join(
  filter(all.oscar.winners, Year < 2016), 
  oscar_matches, by = 'Film')

#Count of missing oscar winners
nrow(missing.oscar.scripts)

#Count of Oscar winners between 1995 and 2015
nrow(filter(all.oscar.winners, Year >= 1995 & Year < 2016))
#1995-2015 oscar winners in Pudding dataset
nrow(filter(oscar_matches, Year >= 1995))

inner_join(missing.oscar.scripts,)
```
This seems like a quite drastic flaw of the Pudding dataset: only 260 of 1271 Oscar winners are present. For the oscar winners between 1995 and 2015 the numbers are somewhat less dramatic: 139 of 303 oscar winners exist in the Pudding dataset.

A potential matching error in the section above is that I only allowed direct name matches. Let's try some fuzzy matching to see if we can find more matches for the Years 1995 to 2015.
```{r}
#Missing oscar winners between 1995 and 2015
missing.oscar.scripts.after.1994 <- 
  filter(missing.oscar.scripts, Year > 1994)

nrow(filter(oscar_matches, Year > 1994))

nrow(missing.oscar.scripts.after.1994)

#install.packages('fuzzyjoin')
library(fuzzyjoin)

library(dplyr)

head(movies)

stringdist_inner_join(missing.oscar.scripts.after.1994, movies,
                      by = c(Film = 'title'))#, max_dist = 4)

inner_join(missing.oscar.scripts.after.1994, movies, 
           by = c(a))
```

```{r}

head(script_summary)

filter(script_summary,)

head(missing.oscar.scripts)

ggplot(data = missing.oscar.scripts, 
       aes(x = Year, color = 'red', alpha = 0.2)) +
  geom_histogram(bins = length(seq(min(missing.oscar.scripts$Year), max(missing.oscar.scripts$Year), 1))) +
  scale_x_continuous(breaks = seq(1920, 2015, 5)) + 
  ggtitle('Missing Oscar winning movies')

View(filter(missing.oscar.scripts, Year > 2000 & Year != 2016))

View(filter(movies, year > 2000))

head(movies)
```

[MASTER LIST]
```{r}
master.list <- read.csv('data/master_list_greatest_films.csv')

head(master.list)

filter(master.list, title == 'Birdman')

View(filter(master.list, grepl('irdman', master.list$title)))

master.matches <- inner_join(script_summary, master.list, by = 'title') 

nrow(master.matches)

tmp_imdb_matches <- inner_join(master.matches, movies, by = 'imdb_id' )

nrow(tmp_imdb_matches)

View(anti_join(master.matches, tmp_imdb_matches, 
          by = c('title' = 'title.x') ))

View(filter(master.matches, script_id == 2380))

filter(movies, script_id == 2380)
filter(master.list, title == 'Get Carter')$imdb_id #== 'tt0208988')

filter(tmp_imdb_matches, imdb_id == 'tt0067128')

head(tmp_imdb_matches)

View(inner_join(master.matches, movies, by = 'title') %>% 
  anti_join(master.matches, by = 'title'))

View(anti_join(master.list, movies, by = 'imdb_id') %>% 
       arrange(rank_ss_2012))#desc(as.numeric(imdb_votes))))

filter(master.matches$, title)

View(head(master.list))

anti_join(inner_join(master.matches, movies, by = 'title'), master.matches, by = 'title')

ggplot(master.list, aes(x = year, alpha = .2)) + 
  geom_histogram(bins = length(seq(min(master.list$year), 
                                   max(master.list$year),1))) +
  geom_histogram(data = master.matches, aes(fill = 'green', color = NULL), 
                 bins = length(seq(min(master.list$year), 
                                   max(master.list$year),1))) + scale_x_continuous(breaks = seq(1920, 2015, 5))
                 
ggplot(master.list, aes(x = as.numeric('year'))) +
  geom_histogram(st)

filter(master.list, grepl('inklater', master.list$director)) %>% 
  inner_join(script_summary, by = 'title')

master.matches[78,]

View(master.matches %>% 
  group_by(director) %>% 
  summarise(mean_words = mean(total_words) ) %>% 
  arrange(desc(mean_words)))
```


[JOINING]
```{r}
#ensuring no duplicate matches
subset(academy.nominated, as.character(Film) == as.character(Name))


inner_join(by = c('title' = 'Film'), #| c('title' = 'Name'), 
           script_summary,academy.nominated)
```

> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}


```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!